Follow-up Samples

### 2017-02-20 submission/manuscript_outline_20161024.Rmd
update fig references in results section

# Lesion Model (Full)
lesion_top10_varsMDA <- read.csv('../results/tables/lesion_model_top_vars_MDA_Summary.csv', header = T)
lesion_imp_vars <- read.csv('../results/tables/rf_otu_tax.csv', header = T, row.names = 1)
lesion_AUC_run_summary <- read.csv('../results/tables/ROC_model_summary.csv', header = T)

# Lesion Model (Reduced)
reduced_lesion_AUC_run_summary <- read.csv('../results/tables/Reduced_ROC_model_summary.csv', header = T)

# IF Model (Full)
IF_top10_varsMDA <- read.csv('../results/tables/IF_model_top_vars_MDA_Summary.csv', header = T)
IF_imp_vars <- read.csv('../results/tables/if_rf_otu_tax.csv', header = T, row.names = 1)
IF_AUC_run_summary <- read.csv('../results/tables/IF_ROC_model_summary.csv', header = T)

# IF Model (Reduced)
reduced_IF_AUC_run_summary <- read.csv('../results/tables/reduced_IF_ROC_model_summary.csv', header = T)

@@

**Results:** The bacterial microbiome in pre and post surgery samples for the adenoma group were more similar to each other than the carcinoma group (P-value `r ifelse(change_summary['thetayc', 'Pvalue'] < 0.05, paste("<"), paste(">"))` 0.05).  There was no difference for either the adenoma or carcinoma group in the relative abundance of any OTU between the pre and post lesion removal (P-value `r ifelse(min(filter(all_OTU_pair_wilcox, analysis == "ALL")[, 'BH_corrected']) < 0.05, paste("<"), paste(">"))` 0.05).  A model built to classify model built to classify lesion had an AUC range of `r format(min(reduced_lesion_AUC_run_summary$ROC), digits = 3)` - `r format(max(reduced_lesion_AUC_run_summary$ROC), digits = 3)` while a model built to classify initial versus follow up samples had an AUC range of `r format(min(reduced_IF_AUC_run_summary$ROC), digits = 3)` - `r format(max(reduced_IF_AUC_run_summary$ROC), digits = 3)`.  The post removal sample for both models had a decrease in the positive probability for either lesion or initial sample (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type == "red_lesion" & comparison =="lesion")[, "BH_correction"] < 0.05, paste("<"), paste(">"))` 0.05).  The lesion model used a total of `r length(lesion_top10_varsMDA$variable)` variables while the initial sample model used a total of `r length(IF_top10_varsMDA$variable)` variables.  A total of `r length(common_vars_summary$otu)` OTUs were common to both models with the majority of these classifying to commensal bacteria (e.g. *`r filter(common_vars_summary, otu == "Otu000028")[, "lowest_ID"]`*, *`r filter(common_vars_summary, otu == "Otu000018")[, "lowest_ID"]`*, *`r filter(common_vars_summary, otu == "Otu000009")[, "lowest_ID"]`*, and *`r filter(common_vars_summary, otu == "Otu000005")[, "lowest_ID"]`*).

@@

***Carcinoma Associated Bacteria:***
Previous literature has suggested that a number of oral microbes may be important in CRC pathogenesis [@flynn_metabolic_2016].  So we next examined whether there were differences in previously well described carcinoma associated OTUs.  These included the OTUs that aligned with *Porphyromonas asaccharolytica* (`r unique(filter(crc_bug_info, tax_id == "Porphyromonas")[, "otu"])`), *Fusobacterium nucleatum* (`r unique(filter(crc_bug_info, tax_id == "Fusobacterium")[, "otu"])`), *Parvimonas micra* (`r unique(filter(crc_bug_info, tax_id == "Parvimonas")[, "otu"])`), and *Peptostreptococcus stomatis* (`r unique(filter(crc_bug_info, tax_id == "Peptostreptococcus")[, "otu"])`).  There was a decrease in relative abundance in initial and follow up samples for lesion and carcinoma for *Parvimonas micra* (P-value `r ifelse(main_crc_bugs_summary["parv", "crc_BH"] < 0.05, paste("<"), paste(">"))` 0.05), and *Porphyromonas asaccharolytica* (P-value `r ifelse(main_crc_bugs_summary["porp", "crc_BH"] < 0.05, paste("<"), paste(">"))` 0.05). In contrast, there was no difference in relative abundance in any of these OTUs for those with adenoma [Figure 2]. We also observed that only a small percentage of those with adenoma or carcinoma were positive or had an appreciable relative abundance of any of these respective OTUs [Figure 2].

***The Lesion Model:***
We next wanted to identify if there were any common bacterial microbiome changes in individuals with adenoma or carcinoma versus normal control individuals.  We investigated this by creating a model to classify lesion versus normal based on the bacterial community and FIT measurements.  This model had an AUC range of `r format(min(reduced_lesion_AUC_run_summary$ROC), digits = 3)` - `r format(max(reduced_lesion_AUC_run_summary$ROC), digits = 3)` after 100 iterations of 20 repeated 10-fold cross validations.  The ROC curve for the final lesion model used was within the observed range of the 100 different test set AUC iterations [Figure 3a].  There were a total of `r length(lesion_top10_varsMDA$variable)` variables [Figure S2].  For this model FIT measurement resulted in the largest decrease in MDA while the OTU with the largest MDA was *`r gsub("_unclassified", "", lesion_imp_vars[as.character((filter(lesion_top10_varsMDA, variable != "fit_result") %>% top_n(1, mean_MDA) %>% select(variable))[, 1]), "Genus"])`* (`r filter(lesion_top10_varsMDA, variable != "fit_result") %>% top_n(1, mean_MDA) %>% select(variable)`) [Figure S2a].

If there were common OTUs that could separate adenoma and carcinoma from normal controls, we would expect to find a decrease in the positive probability of the follow up sample to be a lesion. This is what we observed for the lesion model [Figure 4] (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "lesion")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001).  When we separated individuals based on whether they had an adenoma or carcinoma there was only a decrease in positive probability for the carcinoma group (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "carcinoma_only")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001) and not for the adenoma group (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "all_adenoma")[, "BH_correction"] < 0.05, paste("<"), paste(">"))` 0.05).   We also observed that there were no significant differences in whether the models classified the samples as having lesion between the predicted and actual (P-value `r ifelse(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_lesion" & info == "McnemarPValue")[, "all"] < 0.05, paste("<"), paste(">"))` 0.05).  The lesion model was able to correctly classify the one individual who still had a carcinoma on follow up [Figure 4].

***The Initial Sample Model:***
After building a model to classify based on lesion we wanted to then build a separate model specifically to be able to classify whether samples were initial or follow up samples based on the bacterial community and FIT measurements.  The initial sample model had an AUC range of `r format(min(IF_AUC_run_summary$ROC), digits = 3)` to `r format(max(IF_AUC_run_summary$ROC), digits = 3)` after 100 iterations of 20 repeated 10-fold cross validations.  The test set AUC range for this model performed better then the training set AUCs.  There was a marked decrease in the ROC curve for the final model used when compared to the 100 test set AUC iterations [Figure 3b].  There were a total of `r length(IF_top10_varsMDA$variable)` variables [Figure S2].  For this model *`r gsub("_unclassified", "", IF_imp_vars[as.character((filter(IF_top10_varsMDA, variable != "fit_result") %>% top_n(1, mean_MDA) %>% select(variable))[, 1]), "Genus"])`* (`r filter(IF_top10_varsMDA, variable != "fit_result") %>% top_n(1, mean_MDA) %>% select(variable)`) resulted in the largest decrease in MDA while FIT measurement resulted in the second largest decrease in MDA [Figure S3a].

If there were common OTUs that could separate initial from follow up sample regardless of whether the lesion was adenoma or carcinoma we would expect to find a decrease in the positive probability of the follow up sample to be an initial sample. This is what we observed for the initial sample model [Figure 4] (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_IF" & comparison == "lesion")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001).  When we separated individuals based on whether they had an adenoma or carcinoma there was a decrease in positive probability for both the carcinoma group (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_IF" & comparison == "carcinoma_only")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001) and for the adenoma group (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_IF" & comparison == "all_adenoma")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001).  For this model there was no difference between the predicted and actual classifications (P-value `r ifelse(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_IF" & info == "McnemarPValue")[, "all"] < 0.05, paste("<"), paste(">"))` 0.05) [Figure 4].

>>>>

# Lesion Model 
lesion_top10_varsMDA <- read.csv('../results/tables/reduced_lesion_model_top_vars_MDA_Summary.csv', header = T)
lesion_imp_vars <- read.csv('../results/tables/rf_otu_tax.csv', header = T, row.names = 1)
lesion_AUC_run_summary <- read.csv('../results/tables/Reduced_ROC_model_summary.csv', header = T)

# IF Model
IF_top10_varsMDA <- read.csv('../results/tables/reduced_IF_model_top_vars_MDA_Summary.csv', header = T)
IF_imp_vars <- read.csv('../results/tables/if_rf_otu_tax.csv', header = T, row.names = 1)
IF_AUC_run_summary <- read.csv('../results/tables/reduced_IF_ROC_model_summary.csv', header = T)

@@

**Results:** The bacterial microbiome in pre and post surgery samples for the adenoma group were more similar to each other than the carcinoma group (P-value `r ifelse(change_summary['thetayc', 'Pvalue'] < 0.05, paste("<"), paste(">"))` 0.05).  There was no difference for either the adenoma or carcinoma group in the relative abundance of any OTU between the pre and post lesion removal (P-value `r ifelse(min(filter(all_OTU_pair_wilcox, analysis == "ALL")[, 'BH_corrected']) < 0.05, paste("<"), paste(">"))` 0.05).  A model built to classify model built to classify lesion had an AUC range of `r format(min(lesion_AUC_run_summary$ROC), digits = 3)` - `r format(max(lesion_AUC_run_summary$ROC), digits = 3)` while a model built to classify initial versus follow up samples had an AUC range of `r format(min(IF_AUC_run_summary$ROC), digits = 3)` - `r format(max(IF_AUC_run_summary$ROC), digits = 3)`.  The post removal sample for both models had a decrease in the positive probability for either lesion or initial sample (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type == "red_lesion" & comparison =="lesion")[, "BH_correction"] < 0.05, paste("<"), paste(">"))` 0.05).  The lesion model used a total of `r length(lesion_top10_varsMDA$variable)` variables while the initial sample model used a total of `r length(IF_top10_varsMDA$variable)` variables.  A total of `r length(common_vars_summary$otu)` OTUs were common to both models with the majority of these classifying to commensal bacteria (e.g. *`r filter(common_vars_summary, otu == "Otu000028")[, "lowest_ID"]`*, *`r filter(common_vars_summary, otu == "Otu000018")[, "lowest_ID"]`*, *`r filter(common_vars_summary, otu == "Otu000009")[, "lowest_ID"]`*, and *`r filter(common_vars_summary, otu == "Otu000005")[, "lowest_ID"]`*).

@@

***Carcinoma Associated Bacteria:***
Previous literature has suggested that a number of oral microbes may be important in CRC pathogenesis [@flynn_metabolic_2016].  So we next examined whether there were differences in previously well described carcinoma associated OTUs.  These included the OTUs that aligned with *Porphyromonas asaccharolytica* (`r unique(filter(crc_bug_info, tax_id == "Porphyromonas")[, "otu"])`), *Fusobacterium nucleatum* (`r unique(filter(crc_bug_info, tax_id == "Fusobacterium")[, "otu"])`), *Parvimonas micra* (`r unique(filter(crc_bug_info, tax_id == "Parvimonas")[, "otu"])`), and *Peptostreptococcus stomatis* (`r unique(filter(crc_bug_info, tax_id == "Peptostreptococcus")[, "otu"])`).  There was a decrease in relative abundance in initial and follow up samples for lesion and carcinoma for *Parvimonas micra* (P-value `r ifelse(main_crc_bugs_summary["parv", "crc_BH"] < 0.05, paste("<"), paste(">"))` 0.05), and *Porphyromonas asaccharolytica* (P-value `r ifelse(main_crc_bugs_summary["porp", "crc_BH"] < 0.05, paste("<"), paste(">"))` 0.05) [Figure 2A]. In contrast, there was no difference in relative abundance in any of these OTUs for those with adenoma [Figure 2B]. We also observed that only a small percentage of those with adenoma or carcinoma were positive or had an appreciable relative abundance of any of these respective OTUs [Figure 2].

***The Lesion Model:***
We next wanted to identify if there were any common bacterial microbiome changes in individuals with adenoma or carcinoma versus normal control individuals.  We investigated this by creating a model to classify lesion versus normal based on the bacterial community and FIT measurements.  This model had an AUC range of `r format(min(lesion_AUC_run_summary$ROC), digits = 3)` - `r format(max(lesion_AUC_run_summary$ROC), digits = 3)` after 100 iterations of 20 repeated 10-fold cross validations.  The ROC curve for the final lesion model used was within the observed range of the 100 different test set AUC iterations [Figure 3A].  There were a total of `r length(lesion_top10_varsMDA$variable)` variables [Figure 3B].  For this model FIT measurement resulted in the largest decrease in MDA while the OTU with the largest MDA was *`r gsub("_unclassified", "", lesion_imp_vars[as.character((filter(lesion_top10_varsMDA, variable != "fit_result") %>% top_n(1, mean_MDA) %>% select(variable))[, 1]), "Genus"])`* (`r filter(lesion_top10_varsMDA, variable != "fit_result") %>% top_n(1, mean_MDA) %>% select(variable)`) [Figure 3B].

If there were common OTUs that could separate adenoma and carcinoma from normal controls, we would expect to find a decrease in the positive probability of the follow up sample to be a lesion. This is what we observed for the lesion model (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "lesion")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001).  When we separated individuals based on whether they had an adenoma or carcinoma there was only a decrease in positive probability for the carcinoma group [Figure 3C] (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "carcinoma_only")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001) and not for the adenoma group [Figure 3D] (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "all_adenoma")[, "BH_correction"] < 0.05, paste("<"), paste(">"))` 0.05).   We also observed that there were no significant differences in whether the models classified the samples as having lesion between the predicted and actual (P-value `r ifelse(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_lesion" & info == "McnemarPValue")[, "all"] < 0.05, paste("<"), paste(">"))` 0.05).  The lesion model was able to correctly classify the one individual who still had a carcinoma on follow up [Figure 3C].

***The Initial Sample Model:***
After building a model to classify based on lesion we wanted to then build a separate model specifically to be able to classify whether samples were initial or follow up samples based on the bacterial community and FIT measurements.  The initial sample model had an AUC range of `r format(min(IF_AUC_run_summary$ROC), digits = 3)` to `r format(max(IF_AUC_run_summary$ROC), digits = 3)` after 100 iterations of 20 repeated 10-fold cross validations.  The test set AUC range for this model performed better then the training set AUCs.  There was a marked decrease in the ROC curve for the final model used when compared to the 100 test set AUC iterations [Figure 4A].  There were a total of `r length(IF_top10_varsMDA$variable)` variables [Figure 4B].  For this model *`r gsub("_unclassified", "", IF_imp_vars[as.character((filter(IF_top10_varsMDA, variable != "fit_result") %>% top_n(1, mean_MDA) %>% select(variable))[, 1]), "Genus"])`* (`r filter(IF_top10_varsMDA, variable != "fit_result") %>% top_n(1, mean_MDA) %>% select(variable)`) resulted in the largest decrease in MDA while FIT measurement resulted in the sixth largest decrease in MDA [Figure 4B].

If there were common OTUs that could separate initial from follow up sample regardless of whether the lesion was adenoma or carcinoma we would expect to find a decrease in the positive probability of the follow up sample to be an initial sample. This is what we observed for the initial sample model (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_IF" & comparison == "lesion")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001).  When we separated individuals based on whether they had an adenoma or carcinoma there was a decrease in positive probability for both the carcinoma group [Figure 4C] (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_IF" & comparison == "carcinoma_only")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001) and for the adenoma group [Figure 4D] (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_IF" & comparison == "all_adenoma")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001).  For this model there was no difference between the predicted and actual classifications (P-value `r ifelse(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_IF" & info == "McnemarPValue")[, "all"] < 0.05, paste("<"), paste(">"))` 0.05).

### 2017-02-20 submission/manuscript_outline_20161024.tex
update fig references in results section

\textbf{\emph{Carcinoma Associated Bacteria:}} Previous literature has
suggested that a number of oral microbes may be important in CRC
pathogenesis {[}8{]}. So we next examined whether there were differences
in previously well described carcinoma associated OTUs. These included
the OTUs that aligned with \emph{Porphyromonas asaccharolytica}
(Otu000202), \emph{Fusobacterium nucleatum} (Otu000442),
\emph{Parvimonas micra} (Otu001273), and \emph{Peptostreptococcus
stomatis} (Otu001682). There was a decrease in relative abundance in
initial and follow up samples for lesion and carcinoma for
\emph{Parvimonas micra} (P-value \textless{} 0.05), and
\emph{Porphyromonas asaccharolytica} (P-value \textless{} 0.05). In
contrast, there was no difference in relative abundance in any of these
OTUs for those with adenoma {[}Figure 2{]}. We also observed that only a
small percentage of those with adenoma or carcinoma were positive or had
an appreciable relative abundance of any of these respective OTUs
{[}Figure 2{]}.

\textbf{\emph{The Lesion Model:}} We next wanted to identify if there
were any common bacterial microbiome changes in individuals with adenoma
or carcinoma versus normal control individuals. We investigated this by
creating a model to classify lesion versus normal based on the bacterial
community and FIT measurements. This model had an AUC range of 0.811 -
0.866 after 100 iterations of 20 repeated 10-fold cross validations. The
ROC curve for the final lesion model used was within the observed range
of the 100 different test set AUC iterations {[}Figure 3a{]}. There were
a total of 53 variables {[}Figure S2{]}. For this model FIT measurement
resulted in the largest decrease in MDA while the OTU with the largest
MDA was \emph{Ruminococcaceae} (Otu000251) {[}Figure S2a{]}.

If there were common OTUs that could separate adenoma and carcinoma from
normal controls, we would expect to find a decrease in the positive
probability of the follow up sample to be a lesion. This is what we
observed for the lesion model {[}Figure 4{]} (P-value \textless{}
0.001). When we separated individuals based on whether they had an
adenoma or carcinoma there was only a decrease in positive probability
for the carcinoma group (P-value \textless{} 0.001) and not for the
adenoma group (P-value \textgreater{} 0.05). We also observed that there
were no significant differences in whether the models classified the
samples as having lesion between the predicted and actual (P-value
\textgreater{} 0.05). The lesion model was able to correctly classify
the one individual who still had a carcinoma on follow up {[}Figure
4{]}.

\textbf{\emph{The Initial Sample Model:}} After building a model to
classify based on lesion we wanted to then build a separate model
specifically to be able to classify whether samples were initial or
follow up samples based on the bacterial community and FIT measurements.
The initial sample model had an AUC range of 0.485 to 0.686 after 100
iterations of 20 repeated 10-fold cross validations. The test set AUC
range for this model performed better then the training set AUCs. There
was a marked decrease in the ROC curve for the final model used when
compared to the 100 test set AUC iterations {[}Figure 3b{]}. There were
a total of 70 variables {[}Figure S2{]}. For this model
\emph{Ruminococcaceae} (Otu000278) resulted in the largest decrease in
MDA while FIT measurement resulted in the second largest decrease in MDA
{[}Figure S3a{]}.

If there were common OTUs that could separate initial from follow up
sample regardless of whether the lesion was adenoma or carcinoma we
would expect to find a decrease in the positive probability of the
follow up sample to be an initial sample. This is what we observed for
the initial sample model {[}Figure 4{]} (P-value \textless{} 0.001).
When we separated individuals based on whether they had an adenoma or
carcinoma there was a decrease in positive probability for both the
carcinoma group (P-value \textless{} 0.001) and for the adenoma group
(P-value \textless{} 0.001). For this model there was no difference
between the predicted and actual classifications (P-value \textgreater{}
0.05) {[}Figure 4{]}.

>>>>

\textbf{\emph{Carcinoma Associated Bacteria:}} Previous literature has
suggested that a number of oral microbes may be important in CRC
pathogenesis {[}8{]}. So we next examined whether there were differences
in previously well described carcinoma associated OTUs. These included
the OTUs that aligned with \emph{Porphyromonas asaccharolytica}
(Otu000202), \emph{Fusobacterium nucleatum} (Otu000442),
\emph{Parvimonas micra} (Otu001273), and \emph{Peptostreptococcus
stomatis} (Otu001682). There was a decrease in relative abundance in
initial and follow up samples for lesion and carcinoma for
\emph{Parvimonas micra} (P-value \textless{} 0.05), and
\emph{Porphyromonas asaccharolytica} (P-value \textless{} 0.05)
{[}Figure 2A{]}. In contrast, there was no difference in relative
abundance in any of these OTUs for those with adenoma {[}Figure 2B{]}.
We also observed that only a small percentage of those with adenoma or
carcinoma were positive or had an appreciable relative abundance of any
of these respective OTUs {[}Figure 2{]}.

\textbf{\emph{The Lesion Model:}} We next wanted to identify if there
were any common bacterial microbiome changes in individuals with adenoma
or carcinoma versus normal control individuals. We investigated this by
creating a model to classify lesion versus normal based on the bacterial
community and FIT measurements. This model had an AUC range of 0.811 -
0.866 after 100 iterations of 20 repeated 10-fold cross validations. The
ROC curve for the final lesion model used was within the observed range
of the 100 different test set AUC iterations {[}Figure 3A{]}. There were
a total of 53 variables {[}Figure 3B{]}. For this model FIT measurement
resulted in the largest decrease in MDA while the OTU with the largest
MDA was \emph{Lachnospiraceae} (Otu000015) {[}Figure 3B{]}.

If there were common OTUs that could separate adenoma and carcinoma from
normal controls, we would expect to find a decrease in the positive
probability of the follow up sample to be a lesion. This is what we
observed for the lesion model (P-value \textless{} 0.001). When we
separated individuals based on whether they had an adenoma or carcinoma
there was only a decrease in positive probability for the carcinoma
group {[}Figure 3C{]} (P-value \textless{} 0.001) and not for the
adenoma group {[}Figure 3D{]} (P-value \textgreater{} 0.05). We also
observed that there were no significant differences in whether the
models classified the samples as having lesion between the predicted and
actual (P-value \textgreater{} 0.05). The lesion model was able to
correctly classify the one individual who still had a carcinoma on
follow up {[}Figure 3C{]}.

\textbf{\emph{The Initial Sample Model:}} After building a model to
classify based on lesion we wanted to then build a separate model
specifically to be able to classify whether samples were initial or
follow up samples based on the bacterial community and FIT measurements.
The initial sample model had an AUC range of 0.641 to 0.805 after 100
iterations of 20 repeated 10-fold cross validations. The test set AUC
range for this model performed better then the training set AUCs. There
was a marked decrease in the ROC curve for the final model used when
compared to the 100 test set AUC iterations {[}Figure 4A{]}. There were
a total of 70 variables {[}Figure 4B{]}. For this model
\emph{Pseudomonas} (Otu000438) resulted in the largest decrease in MDA
while FIT measurement resulted in the sixth largest decrease in MDA
{[}Figure 4B{]}.

If there were common OTUs that could separate initial from follow up
sample regardless of whether the lesion was adenoma or carcinoma we
would expect to find a decrease in the positive probability of the
follow up sample to be an initial sample. This is what we observed for
the initial sample model (P-value \textless{} 0.001). When we separated
individuals based on whether they had an adenoma or carcinoma there was
a decrease in positive probability for both the carcinoma group
{[}Figure 4C{]} (P-value \textless{} 0.001) and for the adenoma group
{[}Figure 4D{]} (P-value \textless{} 0.001). For this model there was no
difference between the predicted and actual classifications (P-value
\textgreater{} 0.05).
